---
title: Classification Models (LDA)
author: ~
date: '2019-07-21'
slug: classification-models-lda
categories: 
  - modeling
  - LDA
tags: []
description: ''
featured: ''
featuredalt: ''
featuredpath: ''
linktitle: ''
---
Linear Discriminate Analysis (LDA) is a great classification method.  It is simple and interpretable and can perform as well, if not better, than a lot of other fancier models (neural networks that are not interpretable).


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(MASS)
library(caret)
library(GGally)
```

# Introduction to Linear Discriminant Analysis (LDA)

I will be using the Iris Data set again.  First I will split it into a "training" set and a "test" set.  The test set will represent the unseen observations.

```{r}
iris_data = iris%>%
  mutate(id = row_number())

iris_train = iris_data%>%
  sample_frac(.8)

iris_test = iris_data%>%
  anti_join(iris_train, by = "id")
```

Now I will create an LDA model to classify the species based on the Petal Length and Petal Width.

```{r}
iris_lda = lda(Species ~ Petal.Length + Petal.Width, data = iris_train)

iris_lda

```

```{r}
iris_lda_prediction = predict(iris_lda,iris_test)

table(iris_test$Species,iris_lda_prediction$class)
```

When you look at the approximate decision boundaries, you can see the difference between KNN and LDA.  LDA decision boundaries will be linear.  Figure \@ref(fig:petalbound) shows the decision boundaries.

```{r petalbound, echo=FALSE, fig.cap="Approximate Decision Boundaries for LDA using Petal Length and Petal Width", fig.align="Center"}
iris_lda = lda(Species ~ Petal.Length + Petal.Width, data = iris_train)


pl = seq(min(iris_train$Petal.Length), max(iris_train$Petal.Length), by=0.1)
pw = seq(min(iris_train$Petal.Width), max(iris_train$Petal.Width), by=0.1)

# generates the boundaries for your graph
lgrid <- expand.grid(Petal.Length=pl, 
                     Petal.Width=pw,
                     Sepal.Length = 5.4,
                     Sepal.Width=3.1)

LDAPredGrids <- predict(iris_lda, newdata=lgrid)
LDAPredGrid = as.numeric(LDAPredGrids$class)

iris_train%>%
  ggplot() + 
   geom_point(aes(x=Petal.Length,y=Petal.Width, 
                  color = Species),size=5, shape=1)+
  stat_contour(data = lgrid, aes(x=Petal.Length, y=Petal.Width,
                                 z=LDAPredGrid),bins=2) +
  geom_point(data = lgrid, aes(x=Petal.Length, y=Petal.Width,
                               color=LDAPredGrids$class)) +
  theme_bw()+
  labs(x = "Petal Length", y = "Petal Width")+
  theme(text = element_text(size = 20))

```

## Selecting the proper features

```{r}
iris_lda_sepal = lda(Species ~ Sepal.Length + Sepal.Width,
                     data = iris_train)

iris_lda_sepal
```

```{r}
iris_lda_sepal_prediction = predict(iris_lda_sepal,iris_test)

table(iris_test$Species,iris_lda_sepal_prediction$class)
```

Recall when using KNN the decision boundaries were not at all linear, now with LDA we have linear boundaries.  Figure \@ref(fig:sepalbound) shows the approximate decision boundaries if we were to use sepal length and sepal width.

```{r sepalbound, echo=FALSE, fig.cap="Approximate Decision Boundaries for LDA using Sepal Length and Sepal Width", fig.align="Center"}
sl = seq(min(iris_train$Sepal.Length), max(iris_train$Sepal.Length), by=0.1)
sw = seq(min(iris_train$Sepal.Width), max(iris_train$Sepal.Width), by=0.1)

# generates the boundaries for your graph
lgrid <- expand.grid(Sepal.Length=sl, 
                     Sepal.Width=sw)

LDAPredGrids <- predict(iris_lda_sepal, newdata=lgrid)
LDAPredGrid = as.numeric(LDAPredGrids$class)

iris_train%>%
  ggplot() + 
   geom_point(aes(x=Sepal.Length,y=Sepal.Width, 
                  color = Species),size=5, 
              shape=1)+
  stat_contour(data = lgrid, aes(x=Sepal.Length, y=Sepal.Width,
                                 z=LDAPredGrid),bins=2) +
  geom_point(data = lgrid, aes(x=Sepal.Length, y=Sepal.Width,
                               color=LDAPredGrids$class)) +
  theme_bw()+
  labs(x = "Sepal Length", y = "Sepal Width")+
  theme(text = element_text(size = 20))

```


## Using caret for LDA

```{r}
fitControl = trainControl(
  method = "cv",
  number = 10)

lda_fit = train(Species~Petal.Length+Petal.Width,
                data = iris_train,
                method = "lda",
                trControl = fitControl)

lda_fit
```

```{r}
lda_fit$finalModel
```

